Cudele: A File System with Programmable Consistency and Durability

### 2017-03-27 paper/sections/4_methodology.tex
paper: submit abstract and touch up intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: write first half of results section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Methodology: Decoupled Namespaces}

@@

\begin{table}[t]
\begin{center}
\caption{Future programmers can explore the consistency (C) and
durability (D) spectrums by composing Cudele mechanisms. The consistency
and durability properties are not guaranteed until all mechanisms in the cell
are complete (i.e. the compositions should be considered atomic) and there are
no guarantees while tranisitioning between policies. \label{table:spectrum}}
\begin{tabular}{ l | l | l | l }
  Consistency \(\rightarrow\) &&& \\  
  Durability \(\downarrow\)  & invisible         & eventual        & strong  \\\hline
  none                       & create            & create          & RPCs    \\
                             &                   & +volatile apply &         \\\hdashline
  local                      & create            & create          & RPCs    \\
                             & +local persist    & +local persist  & +local  \\
                             &                   & +volatile apply &  persist\\\hdashline
  global                     & create            & create          & RPCs    \\
                             & +global persist   & +global persist & +stream \\
                             &                   & +volatile apply &         \\
\end{tabular}
\end{center}
\end{table}

>>>>

\section{Methodology: Decoupled Namespaces}
\label{sec:methodology-decoupled-namespaces}

@@

\begin{table}[t]
\begin{center}
\caption{Future programmers can explore the consistency (C) and
durability (D) spectrums by composing Cudele mechanisms. The consistency
and durability properties are not guaranteed until all mechanisms in the cell
are complete (i.e. the compositions should be considered atomic) and there are
no guarantees while tranisitioning between policies. \label{table:spectrum}}
\begin{tabular}{ l | l | l | l }
  C \(\rightarrow\) &&& \\  
  D \(\downarrow\)  & invisible         & eventual        & strong  \\\hline
  none                       & create            & create          & RPCs    \\
                             &                   & +volatile apply &         \\\hdashline
  local                      & create            & create          & RPCs    \\
                             & +local persist    & +local persist  & +local  \\
                             &                   & +volatile apply &  persist\\\hdashline
  global                     & create            & create          & RPCs    \\
                             & +global persist   & +global persist & +stream \\
                             &                   & +volatile apply &         \\
\end{tabular}
\end{center}
\end{table}

### 2017-03-27 paper/sections/6_results.tex
paper: write first half of results section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

This paper adheres to The Popper
Convention\footnote{\url{http://falsifiable.us}}~\cite{jimenez_popper_2016}, so
experiments presented here are available in the repository for this
article\footnote{\url{https://github.com/michaelsevilla/cudele-popper/tree/}}.
Experiments can be examined in more detail, or even re-run, by visiting the
\texttt{[source]} link next to each figure. That link points to a Jupyter
notebook that shows the analysis and source code for that graph and its parent
folder contains all the associated artifacts.

\subsection{Microbenchmarks}
%\begin{figure*}[tb]
%\centering
%\includegraphics[width=180mm]{graphs/slowdown.png}
%\caption{Compared to the \texttt{create} phase, saving and persisting
%updates (\texttt{create+save} and \texttt{create+save+persist}) experience only
%a 4.79\(\times\) and 8.66\(\times\) slowdown, in the worst case for 100K files.
%In contrast, maintaining \texttt{global} consistency is 905.70\(\times\)
%slower.  The disadvantage of decoupling the namespace is the merge phase where
%updates are applied to the metadata store (\texttt{create+apply}, resulting in a
%905.70\(\times\) slowdown for 100K files.}\label{fig:global-v-decoupled}
%\end{figure*}

\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/slowdown-mechanisms.png}
\caption{}
\end{figure}
\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/slowdown-related-work.png}
\caption{}
\end{figure}
\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/slowdown-strong-v-eventual.png}
\caption{}
\end{figure}

\subsubsection{Per phase latencies}

@@

\subsection{Macrobenchmarks}
updatedb: http://lists.ceph.com/pipermail/ceph-users-ceph.com/2015-July/002768.html

>>>>

This paper adheres to The Popper
Convention\footnote{\url{http://falsifiable.us}}~\cite{jimenez_popper_2016}, so
experiments presented here are available in the repository for this
article\footnote{\url{https://github.com/michaelsevilla/cudele-popper/}}.
Experiments can be examined in more detail, or even re-run, by visiting the
\texttt{[source]} link next to each figure. That link points to a Jupyter
notebook that shows the analysis and source code for that graph, which points
to an experiment and its artifacts.

%\begin{figure*}[tb]
%\centering
%\includegraphics[width=180mm]{graphs/slowdown.png}
%\caption{Compared to the \texttt{create} phase, saving and persisting
%updates (\texttt{create+save} and \texttt{create+save+persist}) experience only
%a 4.79\(\times\) and 8.66\(\times\) slowdown, in the worst case for 100K files.
%In contrast, maintaining \texttt{global} consistency is 905.70\(\times\)
%slower.  The disadvantage of decoupling the namespace is the merge phase where
%updates are applied to the metadata store (\texttt{create+apply}, resulting in a
%905.70\(\times\) slowdown for 100K files.}\label{fig:global-v-decoupled}
%\end{figure*}

\subsection{Cudele Mechanism Performance}

\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/slowdown-mechanisms.png}
\caption{
[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/viz-decoupled.ipynb}{source}]
The performance of the Cudele mechanisms normalized to the runtime of the
create mechanism. The runtime of the create mechanism is the time it takes to
write 100 file creates to the client's in-memory journal of metadata
updates.\label{fig:slowdown-mechanisms}}
\end{figure}

\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/behavior-journal-size.png}
\caption{
[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/viz-decoupled.ipynb}{source}]
In-memory client journal scales with the number of
updates.\label{fig:behavior-journal-size}}
\end{figure}

Figure~\ref{fig:slowdown-mechanisms} shows the runtime of the Cudele
mechanisms, normalized to the time it takes to write 100K file create updates
to the client's in-memory journal ({\it i.e.} the create mechanism). Bars above
\(10^0\) are slower than the create mechanism and bars below are faster.
``Stream" is an approximation of the overhead and is calculated by subtracting
the runtime of the job with the journal turned off from the runtime with the
journal turned on. All the slowdowns and speedups reported are for the 100K
file create job, the largest workoad we tested.

% RPCs vs. apply: calls to metadata server vs. RADOS
\subsubsection{Overhead of RPCs} ``RPCs" is \(66\times\) slower than ``volatile
apply" because sending individual metadata updates over the network is costly.
While ``RPCs" sends a request for every file create, ``nonvolatile apply"
writes all the updates to the in-memory journal and applies them to the
in-memory data structures in the metadata server. While communicating the
decoupled namespace directly to the metadata server is faster, communicating
through the object store (``nonvolatile apply") is \(10\times\) slower.

% TODO: why is apply so slow.  
% apply: no CephFS changes, pulls/pushes same RADOS obj.
% v_apply vs. apply/persist: communicating through RADOS
\subsubsection{Overhead of ``nonvolatile apply"} The cost of ``nonvolatile
apply" is much larger than all the other mechanisms.  That mechanism was not
implemented as part of Cudele -- it was a debugging and recovery tool packaged
with CephFS. It works by iterating over the updates in the journal and pulling
all objects that {\it may} be affected by the update.  This means that two
objects are repeatedly pulled, updated, and pushed: the object that houses the
experiment directory and the object that contains the root directory ({\it
i.e.} \texttt{/}).  The cost of communcating through the object store is shown
by comparing the runtime of ``volatile apply" + ``global persist" to
``nonvolatile apply". These two operations end up with the same final metadata
state but using ``nonvolatile apply" is clearly inferior.

% persist vs. save: one disk vs. many
\subsubsection{Parallelism of the Object Store} Comparing ``local" and ``global
persist" demonstrates the bandwidth advantages of storing the journal in a
distributed object store. As the journal size increases, the ``global persist"
performance is \(1.5\times\) faster because the object store is leveraging the
collective bandwidth of the disks in the cluster. This benefit comes from the
object store itself but should be acknowledged when making decisions for the
application; the size of the object store can help mitigate the overheads of
globally persisting metadata updates.

\subsubsection{Journal Size} Figure~\ref{fig:behavior-journal-size} shows the
amount of storage per journal update (\(y\) axis) for the range of file creates
we tested (\(x\) axis). The increase in file size is linear with the number of
metadata creates and suggests that updates for a million files would be
\(2.5\text{KB}*1\text{ million files} = 2.38\text{GB}\). Transfer times for
files this large on an HPC network are reasonable.

\subsection{Eventual vs. Strong Consistency}
\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/slowdown-strong-v-eventual.png}
\caption{The RPC per metadata update of ``Strong Consistency" has a large
overhead compared to the decoupled namespace strategy of ``Eventual
Consistency".\label{fig:slowdown-strong-eventual}}
\end{figure}

Figure~\ref{fig:slowdown-strong-eventual} shows the runtimes of eventual and
strong consistency based systems implemented on Cudele, normalized to the
runtime of the create mechanism (again, just creating files in the client's
in-memory journal). We scaled the number of files up to 100K which is the
maximum size of a directory by default in CephFS. We use the following
compositions from the mechanisms in Table~\ref{table:spectrum}:

\begin{itemize}
  \item Strong Consistency
  \begin{itemize}
    \item[]  \texttt{RPCs} \(+\) \texttt{stream}
  \end{itemize}
  \item Strong Consistency, No Fault Tolerance
  \begin{itemize}
    \item[] \texttt{RPCs}
  \end{itemize}
  \item Eventual Consistency
  \begin{itemize}
    \item[] \texttt{creates} \(+\) \texttt{local persist}
  \end{itemize}
  \item Eventual Consistency, No Fault Tolerance
  \begin{itemize}
    \item[] \texttt{creates} \(+\) \texttt{local persist} \(+\) \texttt{volatile apply}
  \end{itemize}
\end{itemize}

% Final file system metadata states are equivalent No guarantees while
% transitioning mechanisms
We compare these semantics because the final metadata states are equivalent.
Cudele makes no guarantees during execution of the mechanisms or when
transitioning semantics -- the semantics are guaranteed {\it once the mechanism
completes}. So if servers fail during a mechanism, metadata or data may be
lost.

% DeltaFS vs. BatchFS vs. POSIX: decoupled is faster
% (5-7x) vs. (90-104x) slower
% Does not scale?
% Strong Consistency > 10X slower than Eventual Consistency 
\subsubsection{Speedups of Decoupled Namespaces} Eventual consistency uses the
decoupled namespace strategy and shows up to a \(20\times\) speedup over the
traditional namespaces that use RPCs. Compared to the baseline the slowdown is
\(5-7\times\) for Strong Consistency, which emulates BatchFS and
\(90-104\times\) for Eventual Consistency, which emulates DeltaFS.

% POSIX (no stream) vs. POSIX: cost of durability < consistency
\subsubsection{Durability \(<<\) Consistency} The \(1.15\times\) overhead of
``Strong Consistency" compared to ``Strong Consistency, No Fault Tolerance" for
100K files is neglible. It suggests that the overhead of consistency is much
larger than the overhead of durability. This conclusion should be stronger as
we scale the number of files because the cost of streaming the journal into the
object store is constant. We omit the same analysis for ``Eventual Consistency"
because the runtimes are so short that the normalized slowdowns are misleading.

% Eventual Consistency benefit not due to metadata format
\subsubsection{Metadata Formats} Because the metadata formats are the same for
all schemes we argue that the performance gain for decoupled namespaces comes
from relaxing the consistency guarantees and not from the metadata formats, as
was argued in previous work outlining the benefit of
SSTables~\cite{ren:atc2013-tablefs, ren:sc2014-indexfs}.

@@
